---
title: "draft_10_casebase_sim_study"
author: "Trevor Kwan"
date: "13/07/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE}
library(casebase)
library(dotwhisker)
library(tidyr)
library(broom)
library(dplyr)
library(survival)
library(glmnet)
library(ncvreg)
library(stringr)
library(lubridate)
library(knitr)
library(riskRegression)
library(visreg)
library(c060)
```

## Generate Data

- n is # of observation rows
- p is # of predictors
- $\rho$ is the fixed pairwise correlation between any two predictors X_i, X_j
- snr is signal-to-noise ratio

We generated standard Gaussian predictor data, $X$, with $n$ observations and $p$ predictors. We fixed the pairwise correlation between any 2 predictors $X_i$ and $X_j$ at $\rho$, for several values of $\rho$.

$\rho$ is able to fix the correlation between predictors because a larger rho means a larger value is multiplied to the identical columns first (identical columns have corr = 1). When the identical columns later have a standard normal rnorm() added to them, if they are already large values, then they will be resistant to the small addition rnorm() change = more correlated. If they are smaller values, they will be affected a lot by the small addition rnorm() change = less correlated.

We then generated "true" survival times, $Y$, based on $Y = e^{\sum_{j=1}^{p} X_j\beta_j + k*Z}$, where $\beta_j = (-1)^j*e^{-2(j-1)/20}$, $Z \sim N(0,1)$, and $k$ is chosen based on our desired signal-to-noise ratio.

We generated censored times through $C = e^{k*Z}$. The recorded survival time ("d.time") was the minimum of the "true" survival time and censoring time, $T = min{Y, C}$. If the censoring time preceeded the "true" survival time, $C < Y$, the observation was said to be censored.

Regularization Paths for Cox’s Proportional Hazards Model via Coordinate Descent (2011).

```{r}
n = 500
p = 120
rho = 0.5
snr = 3
z = 100
wb = 1 # formula to weight the betas, the larger the wb, the larger the betas
```

```{r}
set.seed(123)

# function to generate x
gen.x = function(n,p,rho){
 if(abs(rho)<1){
  beta=sqrt(rho/(1-rho))
 x0=matrix(rnorm(n*p),ncol=p) # x0 dim is n*p, standard normal (mean = 0, sd = 1) values
 z=rnorm(n)
 x=beta*matrix(z,nrow=n,ncol=p,byrow=F)+x0
 }
 if(abs(rho)==1){ x=matrix(z,nrow=n,ncol=p,byrow=F)}
return(x)
}

# function to generate survival times (introduce sparsity into Beta's)
# z is the number of Beta's you want to artificially set to 0
# p is the total number of Betas/predictors
gen.times = function(x,snr,z){
n=nrow(x)
p=ncol(x)
b=((-1)^(1:p))*exp(-2*( (1:p)-wb)/20)
b[(p-z):p] <- 0
f=x%*%b # x dim is n*p, b dim is p*1, f dim is n*1
e=rnorm(n) # e dim is n*1
k=sqrt(var(f)/(snr*var(e))) # k dim is 1*1, singular value since var of a bunch of numbers is one value
y=exp(f+k*e) # f dim is n*1, k dim is 1*1, e dim is n*1, y dim is n*1
return(y)
}


# function to generate censoring times
gen.times.censor = function(x,snr,z){
n=nrow(x)
p=ncol(x)
b=((-1)^(1:p))*exp(-2*( (1:p)-wb)/20)
b[(p-z):p] <- 0
f=x%*%b
e=rnorm(n)
k=sqrt(var(f)/(snr*var(e)))
y=exp(k*e)
return(y)
}

### testing representation of how rho fixes pairwise correlation
# correlation <- function(x1, x2){
#   sum((x1-mean(x1))*(x2-mean(x2))) / (sqrt(sum((x1-mean(x1))^2) * sum((x2 - mean(x2))^2 )))
# }
# 
# x1 <- 1000*c(1,2,3,4,5) + rnorm(5)
# x2 <- 1000*c(1,2,3,4,5) + rnorm(5)
# correlation(x1, x2)
# 
# x3 <- 1*c(1,2,3,4,5) + rnorm(5)
# x4 <- 1*c(1,2,3,4,5) + rnorm(5)
# correlation(x3, x4)

```

```{r}
set.seed(123)

# generate X's
x_df <- gen.x(n = n, p = p, rho = rho) %>% as.data.frame()

# generate true survival times
surv_times <- gen.times(x = as.matrix(x_df), snr = snr, z = z)

# generate censoring times
cens_times <- gen.times.censor(x = as.matrix(x_df), snr = snr, z = z)

# generate survival/censoring times
y_mat <- cbind(surv_times, cens_times)
d.time <- apply(y_mat, MARGIN = 1, FUN = min)

# generate censoring status (death)
death <- ifelse(cens_times <= surv_times, yes = 0, no = 1)

# combine generated data
df <- cbind(death, d.time, x_df)

# split into train and test: 80% train and 20% test
train_indices <- sample(x = nrow(df), size = 0.80*nrow(df)) # randomly sample 80% of observations
test_indices <- setdiff(x = 1:nrow(df), y = train_indices) # get the remaining 20% of observations

train_df <- df[train_indices,]
test_df <- df[test_indices,]
```

```{r}
# train and test data wrangling

# train data
x_train <- model.matrix(~ . -death -d.time, data = train_df)[, -1] # turns characters into numbers 
y_train <- train_df %>% 
  select(death, d.time) %>%
  data.matrix()

y_train_2 <- with(data = y_train %>% as.data.frame(), expr = Surv(time = d.time, event = death)) # get times and censor status from train_df
# plus/0 is censored

# test data
x_test <- model.matrix(~ . -death -d.time, data = test_df)[, -1]
y_test <- test_df %>% 
  select(death, d.time) %>%
  data.matrix()

y_test_2 <- with(data = y_test %>% as.data.frame(), expr = Surv(time = d.time, event = death))

test_df_model <- cbind(y_test, x_test) %>% as.data.frame()

kable(head(df)[,1:7], caption = "The Generated Data")

# show the true Beta's
b=((-1)^(1:p))*exp(-2*( (1:p)-wb)/20)
b[(p-z):p] <- 0
beta_df <- as.data.frame(round(t(b), 3))
colnames(beta_df) <- c(paste0("B", 1:p))
kable(beta_df, caption = "The Betas")
```


```{r}
# ratio of survival and censored
num_cens <- count(df, death == 0)[2, 2]
num_fail <- count(df, death == 0)[1,2]

num_cens_fail <- as.data.frame(data.frame(num_cens, num_fail) %>% t())
colnames(num_cens_fail) <- c("Number Observation Rows")
rownames(num_cens_fail) <- c("Censored", "Failure")

kable(num_cens_fail, caption = "Ratio of Censored and Failure")
```

# Run Methods

(1) The unpenalized Cox method finds the $\beta$ coefficients that maximizes the likelihood function $L$. 
$L$ is a partial likelihood because it does not consider probabilities for subjects who are censored and it is based on the observed order of events.
$L = L_1 * L_2 * \dots * L_k = \prod_{f=1}^{k}L_f$
$L_f$ measures the hazard of the patient at that time / all subjects at risk at that time. Thus, the baseline hazards will cancel out and we solve for the $\beta's$ that will maximize the product of the $L_f$'s.

(see page 127 of the Kleinbaum book for a worked example.)

Maximizing the log likelihood = minimizing the negative log likelihood.

(2) The elastic net penalized Cox method finds the $\beta$ coefficients that minimizes $\frac{1}{N} * \sum_{i=1}^{N}w_i*l(y_i, \beta_0 + \beta^Tx_i) + \lambda[(1-\alpha)|\beta_2^2|/2+ \alpha|\beta_1|]$.

$l(y_i, \beta_0 + \beta^Tx_i)$ is the negative log likelihood contribution for observation $i$

The elastic net penalty is controlled by $\alpha$, which bridges the gap between lasso ($\alpha = 1$) and ridge ($\alpha = 0$).

The tuning parameter $\lambda$ controls the "strength" or "level" of the penalty.

$w_i$ is for the observation weights, and the default is 1 for each observation.

(3) Casebase is a framework for estimating fully parametric hazard models through logistic regression.

The purpose of the casebase package is to provide practitioners with an easy-to-use software tool to compute a patient's risk (or cumulative incidence) of an event, conditional on a particular patient's covariate profile.

Casebase compares person-moments when the event of interest occurred with person-moments when patients were at risk.

Logistic regression also allows us to perform variable selection within the casebase framework.

```{r}
# fit models

# (1) unpen_cox method
unpen_cox <- coxph(Surv(time = d.time, event = death) ~ ., data = train_df, x = TRUE) # x = TRUE means matrix can be returned in cox_method$x

unpen_cox$coefficients[which(is.na(unpen_cox$coefficients))] <- 0 # set NA coef to 0 for brier score later

# (2) pen_cox method (lasso)
lasso_pen_cox <- cv.glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 1, type.measure = "deviance") # cv does cross validation, in this case it tells you the lambda/model that minimizes the partial likelihood as specified by type.measure = "deviance"
# but the default lambda value is 1SE lambda, must specify lambda.min

# the two dotted lines represent the log of the (1) min lambda and (2) 1 standard error lambda
plot(lasso_pen_cox, type.measure = "deviance")

# (2) pen_cox method (ridge)
ridge_pen_cox <- cv.glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 0, type.measure = "deviance") # cv does cross validation, in this case it tells you the lambda/model that minimizes the partial likelihood as specified by type.measure = "deviance"

# (3) pen_cb method (lasso cb)
lasso_pen_cb <- fitSmoothHazard.fit(x = x_train, y = y_train,
                    family = "glmnet",
                    time = "d.time",
                    event = "death",
                    formula_time = ~ log(d.time), # how hazard depends on time
                    alpha = 1, # lasso
                    ratio = 10, # for each element in your case series, you have this many in your base series
                    standardize = TRUE,
                    penalty.factor = c(0, rep(1, ncol(x_train))) #0 and then number of 1's is number of columns in x
                    )

# (4) pen_cb method (ridge cb)
ridge_pen_cb <- fitSmoothHazard.fit(x = x_train, y = y_train,
                    family = "glmnet",
                    time = "d.time",
                    event = "death",
                    formula_time = ~ log(d.time), # how hazard depends on time
                    alpha = 0, # ridge
                    ratio = 10, # ratio of size of base series to case series
                    standardize = TRUE,
                    penalty.factor = c(0, rep(1, ncol(x_train))) #0 and then number of 1's is number of columns in x
                    )

## testing the link function for logistic regression (visual representation)
# x = seq(from  = 0.01, to = 1, by = 0.01)
# plot(y = log(x/(1-x)), x = x)

## testing to view the casebase series dataframe
# cbSeries <- sampleCaseBase(data = as.data.frame(cbind(y_train, x_train)), time = "d.time", event = "death")
# sampleData <- cbSeries
# sample_event <- as.matrix(sampleData[, "death"]) # get the death col of 0s and 1s from cbSeries
# sample_time <- model.matrix(update(~ log(d.time), ~ . -1), sampleData) # output a column that is the log of d.time in sampleData (only do this if the family = "glmnet" or "gbm") # else, model.matrix(~ log(d.time), sampleData)
# sample_time_x <- cbind(sample_time,
#                        as.matrix(sampleData[, !names(sampleData) %in% c("death", "d.time", "offset")])) # get all the X col and add log(d.time) col
# sample_offset <- sampleData$offset

# what cbSeries looks like before putting it in a logistic regression
# as.data.frame(cbind(sample_event, sample_time_x))
```

# Variable Selection Comparison

True positive: A test result that correctly indicates the presence of a condition or characteristic.
True negative: A test result that correctly indicates the absence of a condition or characteristic.
False positive: A test result which wrongly indicates that a particular condition or attribute is present.
False negative: A test result which wrongly indicates that a particular condition or attribute is absent.

```{r}
# positive = non-zero, negative  = zero
# the higher the TPR and TNR the better, the lowest the FPR and FNR the better
# the range of MCC is from -1 to +1, the higher the MCC the better

# lasso penalized cox variable selection
lasso_cox_vs <- !(as.vector(coefficients(lasso_pen_cox, s = "lambda.min")) == 0) # TRUE if non-zero, FALSE if zero

# true positive (TP), true negative (TN), false positive (FP), false negative (FN)
tp = sum(lasso_cox_vs[1:(p-z)])
tn = sum(!lasso_cox_vs[(p-z+1):p])
fp = z - tn
fn = (p-z)-tp

# true positive rate: TPR = TP/P = TP/(TP+FN) (sensitivity)
tpr_cox = tp/(tp+fn)

# true negative rate: TNR = TN/N = TN/(TN+FP) (specificity)
tnr_cox = tn/(tn+fp)

# false positive rate: (1-TNR)
fpr_cox = 1-tnr_cox

# false negative rate: (1-TPR)
fnr_cox = 1-tpr_cox

# matthew's correlation coefficient (MCC) (TP*TN-FP*FN)/(sqrt((TP+FP)(TP+FN)(TN+FP)(TN+FN)))
mcc_cox = ((tp*tn) - (fp*fn)) / sqrt((tp+fp)*(tp+fn)*(tn+fp)*(tn+fn))

# lasso penalized casebase variable selection
lasso_cb_vs <- !(as.vector(coefficients(lasso_pen_cb, s = "lambda.min")) == 0)

# true positive (TP), true negative (TN), false positive (FP), false negative (FN)
tp = sum(lasso_cb_vs[1:(p-z)])
tn = sum(!lasso_cb_vs[(p-z+1):p])
fp = z - tn
fn = (p-z)-tp

# true positive rate: TPR = TP/P = TP/(TP+FN) (sensitivity)
tpr_cb = tp/(tp+fn)

# true negative rate: TNR = TN/N = TN/(TN+FP) (specificity)
tnr_cb = tn/(tn+fp)

# false positive rate: (1-TNR)
fpr_cb = 1-tnr_cb

# false negative rate: (1-TPR)
fnr_cb = 1-tpr_cb

# matthew's correlation coefficient (MCC) (TP*TN-FP*FN)/(sqrt((TP+FP)(TP+FN)(TN+FP)(TN+FN)))
mcc_cb = ((tp*tn) - (fp*fn)) / sqrt((tp+fp)*(tp+fn)*(tn+fp)*(tn+fn))

# variable selection comparison table
metric_vs <- c("Matthew's Correlation Coefficient (MCC)", "True Positive Rate (TPR)", "True Negative Rate (TNR)", "False Positive Rate (FPR)", "False Negative Rate (FNR)")
method_vs <- c("Lasso Cox", "Lasso Casebase")
values_vs <- round(c(mcc_cox, mcc_cb, tpr_cox, tpr_cb, tnr_cox, tnr_cb, fpr_cox, fpr_cb, fnr_cox, fnr_cb), 3)
vs_df <- matrix(values_vs, ncol = 2, byrow = TRUE) %>% as.data.frame()
colnames(vs_df) <- method_vs
rownames(vs_df) <- metric_vs
kable(vs_df, caption = "Variable Selection Comparison")

df_show <- round(as.data.frame(cbind(as.vector(coefficients(lasso_pen_cox, s = "lambda.min")),as.vector(coefficients(lasso_pen_cb, s = "lambda.min")))),3)
colnames(df_show) <- c("Lasso Cox Betas", "Lasso Casebase Betas")
df_show
```


## Evaluate Metrics

#### 1. Concordance (higher is better)
Concordance is the fraction of all evaluable pairs where the model correctly predicts (using risk score) out of a pairs of individuals.

How to determine if a pair is evaluable?

```{r}
input <- c("censored time", "failure time", "no",
           "failure time", "censored time", "yes",
           "failure time", "failure time", "yes",
           "censored time", "censored time", "no")

eval_df <- matrix(input, ncol = 3, byrow = TRUE) %>% 
  as.data.frame()

colnames(eval_df) <- c("Before", "After", "Evaluable?")

kable(eval_df, caption = "Is the pair evaluable?")
```

Let $X_1$ and $X_2$ represent the risk scores for a given pair, and $T_1$ and $T_2$ the true survival times for a given pair.

For each evaluable pairwise combination pair, there are only 4 possible cases:

1. $X_1 > X_2$ and $T_1 < T_2$ OR $X_1 < X_2$ and $T_1 > T_2$: The pair is concordant (C).
2. $X_1 > X_2$ and $T_1 > T_2$ OR $X_1 < X_2$ and $T_1 < T_2$: The pair is disconcordant (C).
3. $X_1 == X_2$: The risk scores are equal (R).
4. $T_1 == T_2$: The times are equal (T).

The concordance index is:
$$
Concordance = \frac{C + \frac{R}{2}}{C + D + R}
$$


```{r}
# (1) concordance (my function)

# takes in a data frame with columns:
# pred: a column of predicted risk scores from a fitted model using x_test
# time: a column of the true survival time
# status: a column of the true status (censored = 0 or failure = 1)

conc <- function(df){
  row.names(df) <- NULL
  all_pairs <- combn(x = row.names(df), m = 2) %>% as.data.frame()
  
  # get evaluable pairs
  eval <- rep(NA, ncol(all_pairs))
  
  for (i in 1:ncol(all_pairs)) {
    x <- rbind(df[as.numeric(all_pairs[1,i]),],  df[as.numeric(all_pairs[2,i]),])
    if (all(c(1, 1) == x$status)) {
      eval[i] <- "E"
    } else if ((all(c(0, 1) == x$status)) & (x$time[1] > x$time[2])) {
      eval[i] <- "E"
    } else if ((all(c(1, 0) == x$status)) & (x$time[1] < x$time[2])) {
      eval[i] <- "E"
    } else {
      eval[i] <- "I"
    }

  }
  
  # get the class if pairs are evaluable
  class <- rep(NA, ncol(all_pairs))
  
  for (i in 1:ncol(all_pairs)) {
  x <- rbind(df[as.numeric(all_pairs[1,i]),],  df[as.numeric(all_pairs[2,i]),])
  
  if (eval[i] == "E"){
    
  if (((x$pred[1] > x$pred[2]) & (x$time[1] < x$time[2])) | ((x$pred[1] < x$pred[2]) & (x$time[1] > x$time[2]))) {
    class[i] <- "C"
  } else if (((x$pred[1] > x$pred[2]) & (x$time[1] > x$time[2])) | ((x$pred[1] < x$pred[2]) & (x$time[1] < x$time[2]))) {
    class[i] <- "D"
  } else if (x$pred[1] == x$pred[2]) {
    class[i] <- "R"
  } else if (x$time[1] == x$time[2]) {
    class[i] <- "T"
  }
    
  } else {
    class[i] <- "I"
  }
  }
  
  # count the number of C's, D's, and R's and put in C-index formula
  C <- str_count(class, "C") %>% sum()
  D <- str_count(class, "D") %>% sum()
  R <- str_count(class, "R") %>% sum()

  concordance = (C+(R/2))/(C+D+R)
  concordance
}

# test my function using unpen_cox
# pred <- predict(unpen_cox, newdata = as.data.frame(x_test), type = "lp") 
# y_testing <- cbind(y_test[,2], y_test[,1])
# colnames(y_testing) <- c("time", "status")
# 
# df <- as.data.frame(cbind(pred, y_testing))
# conc(df)
```


```{r}
# # (1) concordance with my function
# 
# y_testing <- cbind(y_test[,2], y_test[,1])
# colnames(y_testing) <- c("time", "status")
# 
# # unpen_cox concordance
# unpen_cox_conc <- concordance(unpen_cox)$conc
# # double check concordance function with this
# pred <- predict(unpen_cox, newdata = as.data.frame(x_test), type = "lp") 
# df <- as.data.frame(cbind(pred, y_testing))
# unpen_cox_conc_2 <- conc(df)
# 
# # lasso_pen_cox concordance
# pred <- predict(lasso_pen_cox, newx = x_test, s = "lambda.min") # default is type = link, gives the linear predictors for Cox models
# df <- as.data.frame(cbind(pred, y_testing))
# colnames(df) <- c("pred", "time", "status")
# lasso_cox_conc <- conc(df)
# 
# # ridge_pen_cox concordance
# pred <- predict(ridge_pen_cox, newx = x_test, s = "lambda.min")
# df <- as.data.frame(cbind(pred, y_testing))
# colnames(df) <- c("pred", "time", "status")
# ridge_cox_conc <- conc(df)
# 
# # lasso_pen_cb concordance
# x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
# pred <- predict(lasso_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
# df <- as.data.frame(cbind(pred, y_testing))
# colnames(df) <- c("pred", "time", "status")
# lasso_cb_conc <- conc(df)
# 
# # ridge_pen_cb concordance
# x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
# pred <- predict(ridge_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
# df <- as.data.frame(cbind(pred, y_testing))
# colnames(df) <- c("pred", "time", "status")
# ridge_cb_conc <- conc(df)
# 
# conc_values <- round(rbind(unpen_cox_conc_2, lasso_cox_conc, ridge_cox_conc, lasso_cb_conc, ridge_cb_conc), 4)
# 
# method_names <- c("Unpenalized Cox", "Lasso Penalized Cox", "Ridge Penalized Cox", "Lasso Penalized Casebase", "Ridge Penalized Casebase")
# conc_table <- as.data.frame(cbind(method_names, conc_values))
# rownames(conc_table) <- NULL
# colnames(conc_table) <- c("Method", "Concordance")
# 
# kable(conc_table, caption = "Concordance Table")
```

```{r}
# (1) concordance (C-index method)
y_testing <- cbind(y_test[,2], y_test[,1])
colnames(y_testing) <- c("time", "status")

# unpen_cox concordance
pred <- predict(unpen_cox, newdata = as.data.frame(x_test), type = "lp") 
unpen_cox_conc <- Cindex(pred = pred, y = y_testing)

# lasso_pen_cox concordance
pred <- predict(lasso_pen_cox, newx = x_test, s = "lambda.min")
lasso_cox_conc <- Cindex(pred = pred, y = y_testing)

# ridge_pen_cox concordance
pred <- predict(ridge_pen_cox, newx = x_test, s = "lambda.min")
ridge_cox_conc <- Cindex(pred = pred, y = y_testing)

# lasso_pen_cb concordance
x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
pred <- predict(lasso_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
lasso_cb_conc <- Cindex(pred = pred, y = y_testing)

# ridge_pen_cb concordance
x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
pred <- predict(ridge_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
ridge_cb_conc <- Cindex(pred = pred, y = y_testing)

conc_values <- round(rbind(unpen_cox_conc, lasso_cox_conc, ridge_cox_conc, lasso_cb_conc, ridge_cb_conc), 4)

method_names <- c("Unpenalized Cox", "Lasso Penalized Cox", "Ridge Penalized Cox", "Lasso Penalized Casebase", "Ridge Penalized Casebase")
conc_table <- as.data.frame(cbind(method_names, conc_values))
rownames(conc_table) <- NULL
colnames(conc_table) <- c("Method", "Concordance")

kable(conc_table, caption = "Concordance Table")
```

#### 2. Time-Dependent Brier Scores (lower is better)

The Time-Dependent Brier Scores are computed as $BS(t) = \sum_i\{\frac{I(C_i \ge min(T_i, t))}{\hat{G(min(T_i, t))}}\} \{ I(T_i \le t) - \hat{F}(t|x_i) \}^2$, where $T_i$ represents the event time for observation $i$, $t$ represents the specified time $t$ being evaluated, and $C_i$ represents the censoring indicator or status.

$I(C_i \ge min(T_i, t))$ is an indicator function that evaluates to 0 if the observation is censored and survival time is less than $t$, and evaluates to 1 otherwise.

$\hat{G(min(T_i, t))}$ gives the Kaplan-Meier estimate of the probability of survival for the minimum of the event time and $t$.

$I(T_i \le t)$ is an indicator function that evaluates to 1 if the event time is less than or equal to $t$.

$\hat{F}(t|x_i)$ gives the probability of failure given observation $i$ with covariates $x_i$ at time $t$.

Regularized Regression for Two Phase Failure Time Studies (2021).

```{r}
# # get Brier Score (my method)
# 
# # get_brier_scores function
# get_brier_scores <- function(data, pred) {
#   # set seq of times you want to loop over
#   times <- seq(from = 0,
#                to  = max(data$d.time),
#                length = length(data$d.time))
#   
#   # for each time t, we want a single brier score, summed of the individuals
#   BS_list <- rep(NA, length(times))
#   
#   for (x in 1:length(times)) {
#     t <- times[x]
#     
#     # get the min(d.time, t) of all individuals for a given time t
#     min1 <- sapply(
#       X = data$d.time,
#       FUN = function(x) {
#         min(x, t)
#       }
#     )
#     
#     # get the ind_Ci for all individuals for a given time t
#     ind_Ci <- rep(NA, nrow(data)) # initialize empty vector
#     
#     for (i in 1:nrow(data)) {
#       if ((data$death[i] == 0 & (data$d.time[i] < t)) | (data$death[i] == 1)) {
#         ind_Ci[i] = 1
#       } else {
#         ind_Ci[i] = 0
#       }
#     }
#     
#     # get G(t) estimate for all individuals for a given time t
#     km_fit <- survfit(Surv(d.time, (1 - death)) ~ 1, data = data)
#     
#     G_ti <- summary(km_fit, times = min1)$surv # gives the KM estimate of the probability of survival (censored) for min(T_i, t)
#     
#     G_ti[G_ti == 0] <-
#       0.0001 # add this line so that we don't get NaN when 0/0
#     
#     # get the ind_Ti for all individuals for a given time t
#     ind_Ti <- rep(NA, nrow(data))
#     
#     for (i in 1:nrow(data)) {
#       if (data$d.time[i] <= t) {
#         ind_Ti[i] = 1
#       } else {
#         ind_Ti[i] = 0
#       }
#     }
#     
#     # get the F(t|x_i) for all individuals for a given time t (the probability of failure)
#     F_ti <- pred
#     
#     # get a summed brier score over all individuals for a given time t
#     BS_t <- sum((ind_Ci / G_ti) * ((ind_Ti - F_ti) ^ 2))
#     
#     # add that brier score t to a list
#     BS_list[x] <- BS_t
#     
#   }
#   return(BS_list)
# }
```

```{r}
# # unpen_cox brier score
# pred <- predict(unpen_cox, newdata = as.data.frame(test_df), type = "expected")
# pred <- 1 - exp(-pred)
# Brier <- get_brier_scores(data = test_df, pred = pred)
# times <- seq(from = 0, to  = max(test_df$d.time), length = length(test_df$d.time))
# model <- rep("Unpenalized Cox", length(times))
# unpen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
# colnames(unpen_cox_brier_df) <- c("model", "times", "Brier")
# rownames(unpen_cox_brier_df) <- NULL
# unpen_cox_brier_df$model <- as.factor(unpen_cox_brier_df$model)
# unpen_cox_brier_df$times <- as.double(unpen_cox_brier_df$times)
# unpen_cox_brier_df$Brier <- as.double(unpen_cox_brier_df$Brier)
# 
# # lasso_pen_cox brier score
# y_test_cox <- y_test[,c(2,1)]
# colnames(y_test_cox) <- c("time", "status")
# lasso_pen_cox_2 <- glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 1, lambda = lasso_pen_cox$lambda.min) # here we specify the lambda to be the minimum lambda and refit the glmnet
# lasso_pen_cox_3 <- coxph(Surv(time = d.time, event = death) ~ ., data = train_df, x = TRUE, init = c(as.matrix(coef(lasso_pen_cox_2))), iter.max = 0)# we refit the glmnet as a coxph (should be the same as cv.glmnet object fitted with min lambda)
# # init is the initialized value for where the search begins to find the coef for the model
# # iter.max says to just keep the initial values
# pred <- predict(lasso_pen_cox_3, newdata = as.data.frame(test_df), type = "expected")
# pred <- 1 - exp(-pred)
# Brier <- get_brier_scores(data = test_df, pred = pred)
# times <- seq(from = 0, to  = max(test_df$d.time), length = length(test_df$d.time))
# model <- rep("Lasso Penalized Cox", length(times))
# lasso_pen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
# colnames(lasso_pen_cox_brier_df) <- c("model", "times", "Brier")
# rownames(lasso_pen_cox_brier_df) <- NULL
# lasso_pen_cox_brier_df$model <- as.factor(lasso_pen_cox_brier_df$model)
# lasso_pen_cox_brier_df$times <- as.double(lasso_pen_cox_brier_df$times)
# lasso_pen_cox_brier_df$Brier <- as.double(lasso_pen_cox_brier_df$Brier)
# 
# # ridge_pen_cox brier score
# y_test_cox <- y_test[,c(2,1)]
# colnames(y_test_cox) <- c("time", "status")
# ridge_pen_cox_2 <- glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 0, lambda = ridge_pen_cox$lambda.min)# here we specify the lambda to be the minimum lambda and refit the glmnet
# ridge_pen_cox_3 <- coxph(Surv(time = d.time, event = death) ~ ., data = train_df, x = TRUE, init = c(as.matrix(coef(ridge_pen_cox_2))), iter.max = 0)# we refit the glmnet as a coxph (should be the same as cv.glmnet object fitted with min lambda)
# # init is the initialized value for where the search begins to find the coef for the model
# # iter.max says to just keep the initial values
# pred <- predict(ridge_pen_cox_3, newdata = as.data.frame(test_df), type = "expected")
# pred <- 1 - exp(-pred)
# Brier <- get_brier_scores(data = test_df, pred = pred)
# times <- seq(from = 0, to  = max(test_df$d.time), length = length(test_df$d.time))
# model <- rep("Ridge Penalized Cox", length(times))
# ridge_pen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
# colnames(ridge_pen_cox_brier_df) <- c("model", "times", "Brier")
# rownames(ridge_pen_cox_brier_df) <- NULL
# ridge_pen_cox_brier_df$model <- as.factor(ridge_pen_cox_brier_df$model)
# ridge_pen_cox_brier_df$times <- as.double(ridge_pen_cox_brier_df$times)
# ridge_pen_cox_brier_df$Brier <- as.double(ridge_pen_cox_brier_df$Brier)
# 
# # lasso_pen_cb brier score
# x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
# pred <- predict(lasso_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
# test_df_cb <- as.data.frame(cbind(y_test, x_test_cb))
# colnames(test_df_cb)[3] <- "log(d.time)"
# Brier <- get_brier_scores(data = test_df_cb, pred = pred)
# times <- seq(from = 0, to  = max(test_df$d.time), length = length(test_df$d.time))
# model <- rep("Lasso Penalized Casebase", length(times))
# lasso_pen_cb_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
# colnames(lasso_pen_cb_brier_df) <- c("model", "times", "Brier")
# rownames(lasso_pen_cb_brier_df) <- NULL
# lasso_pen_cb_brier_df$model <- as.factor(lasso_pen_cb_brier_df$model)
# lasso_pen_cb_brier_df$times <- as.double(lasso_pen_cb_brier_df$times)
# lasso_pen_cb_brier_df$Brier <- as.double(lasso_pen_cb_brier_df$Brier)
# 
# # ridge_pen_cb brier score
# x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
# pred <- predict(ridge_pen_cb, newx = x_test_cb, type = "response", s = "lambda.min") # for each subject, this gives the probability that y=1 (death = 1), probability of failure
# test_df_cb <- as.data.frame(cbind(y_test, x_test_cb))
# colnames(test_df_cb)[3] <- "log(d.time)"
# Brier <- get_brier_scores(data = test_df_cb, pred = pred)
# times <- seq(from = 0, to  = max(test_df$d.time), length = length(test_df$d.time))
# model <- rep("Ridge Penalized Casebase", length(times))
# ridge_pen_cb_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
# colnames(ridge_pen_cb_brier_df) <- c("model", "times", "Brier")
# rownames(ridge_pen_cb_brier_df) <- NULL
# ridge_pen_cb_brier_df$model <- as.factor(ridge_pen_cb_brier_df$model)
# ridge_pen_cb_brier_df$times <- as.double(ridge_pen_cb_brier_df$times)
# ridge_pen_cb_brier_df$Brier <- as.double(ridge_pen_cb_brier_df$Brier)
# 
# # combine brier scores
# brier_df <- rbind(unpen_cox_brier_df, lasso_pen_cox_brier_df, ridge_pen_cox_brier_df, lasso_pen_cb_brier_df, ridge_pen_cb_brier_df) %>% 
#   mutate(model = factor(model, levels = c("Unpenalized Cox", "Lasso Penalized Cox", "Ridge Penalized Cox", "Lasso Penalized Casebase", "Ridge Penalized Casebase")))
# 
# # detect outliers
# detect_outlier = function(x) {
#  
#     # calculate first quantile
#     Quantile1 <- quantile(x, probs=.25)
#  
#     # calculate third quantile
#     Quantile3 <- quantile(x, probs=.75)
#  
#     # calculate inter quartile range
#     IQR = Quantile3-Quantile1
#  
#     # return true or false
#     x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
# }
# 
# # subset non-outliers into brier_df
# brier_df <- brier_df[which(!detect_outlier(brier_df$Brier)),]
# 
# # plot brier scores
# ggplot(data = brier_df, aes(x = times, y = Brier, col = model))+
#   geom_line() +
#   xlab("Time") +
#   ylab("Brier Score") +
#   labs(color = "Method", title = "Time-Dependent Brier Scores")
# 
# ## testing logistic regression prob(y=1) formula
# # predict(lasso_pen_cb, newx = x_test_cb[1,], type = "response", s = "lambda.min") # this is prob of y=1
# # b_0 = coefficients(lasso_pen_cb, s = "lambda.min")[1]
# # x_i = as.vector(x_test_cb[1,])
# # betas = as.vector(coefficients(lasso_pen_cb, s = "lambda.min"))[-c(1)]
# # (exp(b_0 + betas%*%x_i)/(1 + exp(b_0 + betas%*%x_i))) # this is prob of y=1 for logistic regression
# # 1/(1 + exp(-(b_0+x_i%*%betas))) # this is also prob of y=1 for logistic regression but simplified
```

```{r}
# Absolute Risks
newx <- model.matrix(death ~ . - d.time,
                     data = test_df)[, -c(1)] # get all x covariate values from test dataset
times <- sort(unique(test_df$d.time)) # get all unique survival times (d.time) and sort in ascending order

# 1. Unpenalized Cox
ab_unpen_cox <- survival::survfit(unpen_cox, newdata = test_df)


# 2. Lasso Penalized Cox
y_test_cox <- y_test[,c(2,1)]
colnames(y_test_cox) <- c("time", "status")
lasso_pen_cox_2 <- glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 1, lambda = lasso_pen_cox$lambda.min) # here we specify the lambda to be the minimum lambda and refit the glmnet
nonzero_covariate_lasso_cox <- predict(lasso_pen_cox_2, type = "nonzero", s = "lambda.min") # get the columns of the non-zero selected covariates
nonzero_coef_cox <- coef(lasso_pen_cox_2)
# creating a new dataset containing the chosen covariates with predict nonzero
cleanCoxData <- as.data.frame(cbind(y_train, x_train[,nonzero_covariate_lasso_cox[,1]]))

lasso_pen_cox_4 <- coxph(Surv(time = d.time, event = death) ~ ., data = cleanCoxData, x = TRUE)# we refit the glmnet as a coxph using the "cleaned" data which includes only nonzero selected beta columns from lasso cox

newdata_lasso_pen_cox <- as.data.frame(newx[, nonzero_covariate_lasso_cox[,1]])
colnames(newdata_lasso_pen_cox) <- colnames(cleanCoxData)[-c(1,2)]

ab_lasso_cox <- survival::survfit(lasso_pen_cox_4, type = "breslow", 
                              newdata = newdata_lasso_pen_cox)

# 3. Ridge Penalized Cox
y_test_cox <- y_test[,c(2,1)]
colnames(y_test_cox) <- c("time", "status")
ridge_pen_cox_2 <- glmnet(x = x_train, y = y_train_2, family = "cox", alpha = 0, lambda = ridge_pen_cox$lambda.min)# here we specify the lambda to be the minimum lambda and refit the glmnet
ridge_pen_cox_3 <- coxph(Surv(time = d.time, event = death) ~ ., data = train_df, x = TRUE, init = c(as.matrix(coef(ridge_pen_cox_2))), iter.max = 0)# we refit the glmnet as a coxph (should be the same as cv.glmnet object fitted with min lambda)

ab_ridge_cox <- survival::survfit(ridge_pen_cox_3, type = "breslow", 
                              newdata = as.data.frame(newx))

# 4. Lasso Penalized Case-base
ab_lasso_pen_cb <- casebase::absoluteRisk(lasso_pen_cb,
  time = times,
  newdata = newx,
  s = "lambda.1se",
  method = "numerical"
)

# 5. Ridge Penalized Case-base
ab_ridge_pen_cb <- casebase::absoluteRisk(ridge_pen_cb,
  time = times,
  newdata = newx,
  s = "lambda.1se",
  method = "numerical"
)

# option to filter some observations to make it more stepwise for Cox (didn't use this)
cox_ind <- seq(from = 1, to = length(ab_unpen_cox$time), 1)

# Combine absolute risk estimates
# for each time t, take the average of probs for each individual (1 minus for the prob of failure)
data_absRisk <- bind_rows(
  data.frame(Time = ab_unpen_cox$time[cox_ind], Prob = 1 - rowMeans(ab_unpen_cox$surv[cox_ind,]),
             Model = "Unpenalized Cox"),
  data.frame(Time = ab_lasso_cox$time[cox_ind], Prob = 1 - rowMeans(ab_lasso_cox$surv[cox_ind,]),
             Model = "Lasso Penalized Cox"),
    data.frame(Time = ab_ridge_cox$time[cox_ind], Prob = 1 - rowMeans(ab_ridge_cox$surv[cox_ind,]),
             Model = "Ridge Penalized Cox"),
  data.frame(Time = ab_lasso_pen_cb[, 1], Prob = rowMeans(ab_lasso_pen_cb[, -c(1)]),
             Model = "Lasso Penalized Casebase"),
  data.frame(Time = ab_ridge_pen_cb[, 1], Prob = rowMeans(ab_ridge_pen_cb[, -c(1)]),
             Model = "Ridge Penalized Casebase")
)

############ testing work CI
# ab_unpen_cox$time %>% length()
# ab_lasso_pen_cb[, 1] %>% length()
# 
# rowMeans(ab_unpen_cox$surv) %>% length()
# rowMeans(ab_lasso_pen_cb[, -c(1)]) %>% length()
# 
# data_absRisk %>% filter(Model == "Lasso Penalized Casebase") %>% select(c("Prob")) %>% unique()
# data_absRisk %>% filter(Model == "Lasso Penalized Cox") %>% select(c("Prob")) %>% unique()
# 
# data_absRisk
```

```{r absRiskPlot, echo=FALSE, eval=eval_cs3}
# helper functions scale
scaleFUNy <- function(x) sprintf("%.2f", x)
scaleFUNx <- function(x) sprintf("%.0f", x)

# detect outliers
# detect_outlier = function(x) {
# 
#     # calculate first quantile
#     Quantile1 <- quantile(x, probs=.25)
# 
#     # calculate third quantile
#     Quantile3 <- quantile(x, probs=.75)
# 
#     # calculate inter quartile range
#     IQR = Quantile3-Quantile1
# 
#     # return true or false
#     x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
# }

# subset non-outliers into brier_df
# data_absRisk <- data_absRisk[which(!detect_outlier(data_absRisk$Time)),]

reg_ci <- ggplot(data_absRisk, aes(Time, Prob)) +
  geom_line(aes(colour = Model)) +
  labs(y = "Probability of failure", x = "Time", color = "Method", title = "Cumulative Incidence Curves") +
  expand_limits(y = 1) +
  scale_y_continuous(labels = scaleFUNy, n.breaks = 9) +
  xlim(c(0,max(filter(data_absRisk, Model == "Lasso Penalized Casebase")$Time))) # set a limit here because there can be some outlier times

reg_ci
```


```{r}
# get Brier Score (my method)

# get_brier_scores function
get_brier_scores <- function(data, pred, times) {

  # for each time t, we want a single brier score, summed of the individuals
  BS_list <- rep(NA, length(times))
  
  for (x in 1:length(times)) {
    t <- times[x]
    
    # get the min(d.time, t) of all individuals for a given time t
    min1 <- sapply(
      X = data$d.time,
      FUN = function(x) {
        min(x, t)
      }
    )
    
    # get the ind_Ci for all individuals for a given time t
    ind_Ci <- rep(NA, nrow(data)) # initialize empty vector
    
    for (i in 1:nrow(data)) {
      if ((data$death[i] == 0 & (data$d.time[i] < t)) | (data$death[i] == 1)) {
        ind_Ci[i] = 1
      } else {
        ind_Ci[i] = 0
      }
    }
    
    # get G(t) estimate for all individuals for a given min(t, T_i)
    km_fit <- survfit(Surv(d.time, (1 - death)) ~ 1, data = data)
    
    G_ti <- summary(km_fit, times = min1)$surv # gives the KM estimate of the probability of survival (censored) for min(T_i, t)
    
    G_ti[G_ti == 0] <-
      0.0001 # add this line so that we don't get NaN when 0/0
    
    # get the ind_Ti for all individuals for a given time t
    ind_Ti <- rep(NA, nrow(data))
    
    for (i in 1:nrow(data)) {
      if (data$d.time[i] <= t) {
        ind_Ti[i] = 1
      } else {
        ind_Ti[i] = 0
      }
    }

    # get the F(t|x_i) for all individuals for a given time t (the probability of failure)
    F_ti <- pred[x,]

    # get a summed brier score over all individuals for a given time t
    BS_t <- sum((ind_Ci / G_ti) * ((ind_Ti - F_ti) ^ 2))

    # add that brier score t to a list
    BS_list[x] <- BS_t
    
  }
  return(BS_list)
}
```



```{r}
# unpen_cox brier score
Brier <- get_brier_scores(data = test_df, pred = (1 - ab_unpen_cox$surv), times = ab_unpen_cox$time)
times <- ab_unpen_cox$time
model <- rep("Unpenalized Cox", length(times))
unpen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
colnames(unpen_cox_brier_df) <- c("model", "times", "Brier")
rownames(unpen_cox_brier_df) <- NULL
unpen_cox_brier_df$model <- as.factor(unpen_cox_brier_df$model)
unpen_cox_brier_df$times <- as.double(unpen_cox_brier_df$times)
unpen_cox_brier_df$Brier <- as.double(unpen_cox_brier_df$Brier)

# lasso_pen_cox brier score
Brier <- get_brier_scores(data = test_df, pred = (1 - ab_lasso_cox$surv), times = ab_lasso_cox$time)
times <- ab_lasso_cox$time
model <- rep("Lasso Penalized Cox", length(times))
lasso_pen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
colnames(lasso_pen_cox_brier_df) <- c("model", "times", "Brier")
rownames(lasso_pen_cox_brier_df) <- NULL
lasso_pen_cox_brier_df$model <- as.factor(lasso_pen_cox_brier_df$model)
lasso_pen_cox_brier_df$times <- as.double(lasso_pen_cox_brier_df$times)
lasso_pen_cox_brier_df$Brier <- as.double(lasso_pen_cox_brier_df$Brier)

# ridge_pen_cox brier score
Brier <- get_brier_scores(data = test_df, pred = (1 - ab_ridge_cox$surv), times = ab_lasso_cox$time)
times <- ab_lasso_cox$time
model <- rep("Ridge Penalized Cox", length(times))
ridge_pen_cox_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
colnames(ridge_pen_cox_brier_df) <- c("model", "times", "Brier")
rownames(ridge_pen_cox_brier_df) <- NULL
ridge_pen_cox_brier_df$model <- as.factor(ridge_pen_cox_brier_df$model)
ridge_pen_cox_brier_df$times <- as.double(ridge_pen_cox_brier_df$times)
ridge_pen_cox_brier_df$Brier <- as.double(ridge_pen_cox_brier_df$Brier)

# lasso_pen_cb brier score
x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
test_df_cb <- as.data.frame(cbind(y_test, x_test_cb))
colnames(test_df_cb)[3] <- "log(d.time)"
Brier <- get_brier_scores(data = test_df_cb, pred = ab_lasso_pen_cb[, -c(1)], times = ab_lasso_pen_cb[, 1])
times <- ab_lasso_pen_cb[, 1]
model <- rep("Lasso Penalized Casebase", length(times))
lasso_pen_cb_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
colnames(lasso_pen_cb_brier_df) <- c("model", "times", "Brier")
rownames(lasso_pen_cb_brier_df) <- NULL
lasso_pen_cb_brier_df$model <- as.factor(lasso_pen_cb_brier_df$model)
lasso_pen_cb_brier_df$times <- as.double(lasso_pen_cb_brier_df$times)
lasso_pen_cb_brier_df$Brier <- as.double(lasso_pen_cb_brier_df$Brier)

# ridge_pen_cb brier score
x_test_cb <- cbind(log(y_test[,"d.time"]), x_test) # solution: add log(d.time) as a first column to x_test
test_df_cb <- as.data.frame(cbind(y_test, x_test_cb))
colnames(test_df_cb)[3] <- "log(d.time)"
Brier <- get_brier_scores(data = test_df_cb, pred = ab_ridge_pen_cb[, -c(1)], times = ab_ridge_pen_cb[,1])
times <- ab_ridge_pen_cb[,1]
model <- rep("Ridge Penalized Casebase", length(times))
ridge_pen_cb_brier_df <- as.data.frame(cbind(model, times, Brier))[order(times),]
colnames(ridge_pen_cb_brier_df) <- c("model", "times", "Brier")
rownames(ridge_pen_cb_brier_df) <- NULL
ridge_pen_cb_brier_df$model <- as.factor(ridge_pen_cb_brier_df$model)
ridge_pen_cb_brier_df$times <- as.double(ridge_pen_cb_brier_df$times)
ridge_pen_cb_brier_df$Brier <- as.double(ridge_pen_cb_brier_df$Brier)


# combine brier scores
brier_df <- rbind(unpen_cox_brier_df, lasso_pen_cox_brier_df, ridge_pen_cox_brier_df, lasso_pen_cb_brier_df, ridge_pen_cb_brier_df) %>% 
  mutate(model = factor(model, levels = c("Unpenalized Cox", "Lasso Penalized Cox", "Ridge Penalized Cox", "Lasso Penalized Casebase", "Ridge Penalized Casebase")))

# detect outliers
detect_outlier = function(x) {
 
    # calculate first quantile
    Quantile1 <- quantile(x, probs=.25)
 
    # calculate third quantile
    Quantile3 <- quantile(x, probs=.75)
 
    # calculate inter quartile range
    IQR = Quantile3-Quantile1
 
    # return true or false
    x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
}

# subset non-outliers into brier_df
brier_df <- brier_df[which(!detect_outlier(brier_df$Brier)),]

# plot brier scores
ggplot(data = brier_df, aes(x = times, y = Brier, col = model))+
  geom_line() +
  xlab("Time") +
  ylab("Brier Score") +
  labs(color = "Method", title = "Time-Dependent Brier Scores") +
  xlim(c(0,max(filter(brier_df, model == "Ridge Penalized Casebase")$times))) # set a limit here because there can be some outlier times
```

```{r}
# ########### get brier score from riskregression
# # brierCoxKM <- Score(list("Cox" = cox,
# #                          "K-M" = KM), data = test,
# #                    formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
# #                    se.fit = FALSE, metrics = "brier", contrasts = FALSE,
# #                    times = times)
# # # 2. Penalized models
# # brierPenalized <- Score(list("Pen. Cox" = coxNet,
# #                              "Pen. CB" = pen_cb),
# #                         data = cbind(subset(test, select = c(d.time, death)),
# #                                      as.data.frame(newx)),
# #                         formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
# #                         se.fit = FALSE, metrics = "brier", contrasts = FALSE,
# #                         times = times)
# 
# times <- sort(unique(test_df$d.time))
# 
# brier_unpen_cox <- Score(list("Unpenalized Cox" = unpen_cox), data = as.data.frame(test_df),
#      formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
#      se.fit = FALSE, metrics = "brier", contrasts = FALSE,
#      times = times)
# 
# brier_lasso_cox <- Score(list("Lasso Cox" = lasso_pen_cox_4), data = as.data.frame(test_df),
#      formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
#      se.fit = FALSE, metrics = "brier", contrasts = FALSE,
#      times = times)
# 
# brier_ridge_cox <- Score(list("Ridge Cox" = ridge_pen_cox_3), data = as.data.frame(test_df),
#      formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
#      se.fit = FALSE, metrics = "brier", contrasts = FALSE,
#      times = times)
# 
# brier_lasso_cb <- Score(list("Lasso Casebase" = lasso_pen_cb), data = as.data.frame(test_df),
#      formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
#      se.fit = FALSE, metrics = "brier", contrasts = FALSE,
#      times = times)
# 
# brier_ridge_cb <- Score(list("Ridge Casebase" = ridge_pen_cb), data = as.data.frame(test_df),
#      formula = Hist(d.time, death != 0) ~ 1, summary = NULL,
#      se.fit = FALSE, metrics = "brier", contrasts = FALSE,
#      times = times)
# 
# 
# # Combine scores
# data_brier <- bind_rows(
#   brier_unpen_cox$Brier$score %>% 
#     mutate(model = as.character(model)) %>% 
#     filter(model != "Null model"),
#   
#     brier_lasso_cox$Brier$score %>% 
#     mutate(model = as.character(model)) %>% 
#     filter(model != "Null model"),
#   
#     brier_ridge_cox$Brier$score %>% 
#     mutate(model = as.character(model)) %>% 
#     filter(model != "Null model"),
#   
#     brier_lasso_cb$Brier$score %>% 
#     mutate(model = as.character(model)) %>% 
#     filter(model != "Null model"),
#   
#     brier_ridge_cb$Brier$score %>% 
#     mutate(model = as.character(model)) %>% 
#     filter(model != "Null model")
#   ) %>% 
#   mutate(model = factor(model, levels = c("Unpenalized Cox", "Lasso Cox", "Ridge Cox", "Lasso Casebase", "Ridge Casebase")))
# 
# # remove na's
# data_brier <- na.omit(data_brier)
# 
# # detect outliers
# detect_outlier = function(x) {
#  
#     # calculate first quantile
#     Quantile1 <- quantile(x, probs=.25)
#  
#     # calculate third quantile
#     Quantile3 <- quantile(x, probs=.75)
#  
#     # calculate inter quartile range
#     IQR = Quantile3-Quantile1
#  
#     # return true or false
#     x > Quantile3 + (IQR*1.5) | x < Quantile1 - (IQR*1.5)
# }
# 
# # subset non-outliers into brier_df
# data_brier <- data_brier[which(!detect_outlier(data_brier$times)),]
# 
# reg_brier <- ggplot(data = data_brier, aes(x = times, y = Brier, col = model)) +
#   geom_line() +
#   xlab("Time") +
#   ylab("Brier Score") +
#   labs(color = "Method", title = "Time-Dependent Brier Scores") +
#   scale_x_continuous(labels = scaleFUNx, 
#                      breaks = round(seq(0, 5, by = 1))) + 
#   scale_y_continuous(labels = scaleFUNy, n.breaks = 8)
# reg_brier
```

## PT Graphs

```{r}
pt_df <- df[which(!detect_outlier(df$d.time)),]
pt_object <- popTime(data = pt_df, event = "death", time = "d.time")
plot(pt_object, xlab = "Time", add.case.series = FALSE)

```


```{r}
plot(pt_object, xlab = "Time", )


```

```{r}
plot(pt_object,
     add.case.series = TRUE,
     add.base.series = TRUE,
     comprisk = FALSE,
     xlab = "Time")

```












